A 30 Million Year Old Mini-Neptune in the Kepler Field

### 2021-05-27 paper/ms.tex
wow, nature papers are short

\begin{abstract}
  Kepler\,1627 is an active Sun-like star previously found to host a
  $3.56\pm 0.04\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  selected for observation by the Kepler satellite because it is
  nearby and Sun-like.  Here, combining data from the Gaia satellite
  with rotation periods from TESS, we show that Kepler\,1627 is a
  member of the underappreciated \cn.  Based on isochronal and
  gyrochronal evidence, the cluster is $30\pm
  XX$\,Myr old.  Based on archival NIRC2 imaging, we also highlight the existence of 
  a previously unknown stellar companion. Our analysis advances Kepler\,1627Ab as the youngest planet with a
  precise age observed by the main Kepler mission.  Newly
  derived ages from Gaia offer the opportunity to significantly expand
  the census of age-dated planets -- we offer a literature compilation
  of young stars to support this effort.  The properties of \pn\
  may also help clarify how the orbits and atmospheres of the
  mini-Neptune planets evolve.
\end{abstract}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf 100 day window of the \sn\ light curve.}  
		The full Kepler dataset spans 1{,}437 days (3.9 years), sampled at
		30 minute cadence.  The {\it top panel} shows the
		\texttt{PDCSAP} median-subtracted flux in units of
		parts-per-thousand ($\times 10^{-3}$).  The dominant signal
		is induced by starspots and plages.  The model for the
		stellar variability (orange line) is subtracted in the {\it
			bottom panel}, revealing the transits of \pn, as well as other
		deviations from the stellar variability model.  The 
		Figure Set available online shows the entire 3.9 years of
		available data.
		\label{fig:lightcurve}
	}
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f5.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Phase-folded transit of Kepler 1627b, with stellar
			variability removed.}  
		The 1-$\sigma$ model uncertainties and the mean model are shown
		as the faint purple band, and the dark purple line.
		Gray points are individual \texttt{PDCSAP} flux measurements;
		black points are binned to 20 minutes, with a representative error
		bar shown.
		{\bf LGB FIXME: seems like the uncertainty is bigger than the RMS
			in the out of transit...}
		\label{fig:phasefold}
	}
\end{figure*}

@@

Comparing our young star list against the Kepler field yielded two
notable discoveries.  The first, to be discussed in an upcoming analysis
by J. Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr
open cluster Theia\,520 (UBC\,1).  At least five Kepler planets are
known in the cluster, including the Kepler-52 and Kepler-968 systems.
The second discovery, and the focus of this work, is that Kepler
observed a portion of the $\delta$\,Lyr cluster, shown in
Figure~\ref{fig:skychart}.

While we have yet to perform an in-depth search of the {\bf XXX} light
curves Kepler acquired for the \cn, there is already a known planet
that has been detected in the cluster: Kepler\,1627b.  A portion of
the light curve is shown in Figure~\ref{fig:lightcurve}, and the
phase-folded transit is shown in Figure~\ref{fig:phasefold}.  The
dominant signal in the \texttt{PDCSAP} photometry is a quasi-period
starspot signal with a peak-to-peak amplitude that varies between
roughly 2\% and 8\%, depending on the degree of asymmetry in spot
coverage between the stellar hemispheres.  The starspot signal repeats
roughly every 2.606\,days, though this value varies by a few percent
depending on which portion of the light curve is analyzed {\bf (FIXME:
	do!)}.  The details of how we model this signal to derive the
parameters of the planet are discussed in Section~\ref{app:gptransit}.
The results, given in Table~\ref{tab:posterior}, are consistent with a
mini-Neptune sized planet ($3.56\pm 0.04\,R_\oplus$) on a circular
7.2\,day orbit, around a G8V star ($0.91 \pm 0.05 R_\odot$) {\bf
	fixme: age}.

Could the transit signal be produced by anything other than a planet
orbiting this near-solar analog?  \citet{morton_false_2016} considered
the various false positive scenarios, including those of a foreground
eclipsing binary, a hierarchical triple-star binary system, and a
background eclipsing binary.  \citet{morton_false_2016} reported the latter
scenario to be the only possible case, with a (model-depending)
probability of $\approx10^{-5}$.  While this statistical validation is
reassuring, ``validated" planets have been demonstrated to be false
positives \citep{shporer_three_2017}.  We are not particularly
concerned in this case however, since these cases almost always
originate when attempting to separate the smallest stars from
identically-sized gas giant planets.  With a size of $\approx
3.6\,R_\oplus$, Kepler\,1627b is well below the regime of concern.

Nonetheless, our analysis of archival Keck-NIRC2 Kp-band (2.12\,$\mu
$m) imaging did reveal the existence of a neighboring star, unresolved
in the Gaia source catalog {\bf FIXME: uncertainties}.  The NIRC2
images yield a projected separation $\rho=0\farcs17$, with $\Delta
{\rm Kp} = 2.5$.  Using the measured Gaia EDR3 parallax for the
system, this implies a projected separation of 54\,AU.  The presence
of this star is consistent with the excess noise reported to be
present in the Gaia astrometric time-series (see
Section~\ref{app:companionstar}).  Given the low chance of a star
imaged within {\bf X.X arcseconds} to be a chance unbound companion
along this line of sight, we dub the companion star Kepler\,1627B, and
proceed under the assumption that it is bound.

Unforuntately, we do not have any reliable color information about the
secondary star.  Based on the
tabulation\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
version \texttt{2021/03/02}.} by \citet{pecaut_mamajek_2013}, the
NIR-contrast measured for a main sequence G8V star corresponds to a
spectral type for the companion of $\approx$M2V ($M_\star \approx
0.44\,M_\odot$).  However, the companion should therefore have a
significantly longer pre-main-sequence contraction phase, which would
imply that this mass is an overestimate by $\approx20\%$.

Could the companion host the planet, or be creating a false positive
signal?  The star contributes $\approx$1\% of the total flux observed
in the Kepler aperture(0.7-2.0\% V-band to G-band dmag difference from
EEM table; todo is fix using model spectra).  The observed transit has
a depth of 0.18\%.  An 18\% deep eclipse of the secondary star could
produce such a signal.  Given the shape of the observed signal
(allowed impact parameters span 0.02 to 0.73, see
Table~\ref{tab:posterior}), the body transiting the secondary would
need to be non-grazing with $R/R_2 \approx 0.42$, {\it i.e.}, with
radius $\approx 0.2R_\odot$.
The duration of the transit 
($2.823 \pm 0.057 hr$) is ultimately what makes this scenario
implausible.

% {\bf TODO:
% uhhh what about the pre-MS M dwarf star being not very dense?
% What is the density of the G dwarf expected to be?
% }
% 
% $T_{14}$ & hr & -- & 2.822 & 2.823 & 0.057 & 2.710 & 2.917 & 1086.008 & 0.002 \\
% $\rho_\star$ & g$\ $cm$^{-3}$ & -- & 2.27 & 2.312 & 0.516 & 1.408 & 3.272 & 910.746 & 0.001 \\

%\section{Discussion}
%\label{sec:discussion}

%\section{Conclusion}
%\label{sec:conclusion}

@@

\section{Companion Star}
\label{app:companionstar}

{\bf TODO: include the RUWE vs Rp mag.}

>>>>

\begin{abstract}
  Kepler\,1627 is an active G8V star previously found to host a
  $3.56\pm 0.04\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star
  was selected for observation by the Kepler satellite because it is
  nearby and it resembles the Sun.  Combining data from the Gaia
  satellite with rotation periods from TESS, we show that Kepler\,1627
  is a member of the \cn.  Based on isochronal and gyrochronal
  evidence, the cluster is $30\pm XX$\,Myr old.  Based on archival
  NIRC2 imaging, we also note the existence of a low-mass stellar
  companion. Our analysis advances Kepler\,1627Ab as the youngest
  planet with a precise age detected by the main Kepler mission.
  Newly derived ages from Gaia offer the opportunity to significantly
  expand the census of age-dated planets -- we offer a literature
  concatenation of young stars to support this effort.  The properties
  of \pn\ may also help clarify how the orbits and atmospheres of the
  mini-Neptune planets evolve.
\end{abstract}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{TEMP_Prot_vs_BpmRp.png}
			\includegraphics[width=0.49\textwidth]{TEMP_hrd.png}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf \cn is $35\pm15$\,Myr old.} {\it Left:} TESS and
    Kepler rotations periods and dereddened Gaia colors, with the
    Pleiades \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
    \citep[650\,Myr;][]{douglas_poking_2017} shown for reference.
    Stars of a given color (mass) move up through time due to magnetic
    braking.  The \cn\ has a 40 Myr gyrochronal age, as calibrated
    against the LDB-age of IC\,2602.  {\it Right:} HR diagram of the
    same comparison. {\bf todo: caption improve}.
   \label{fig:age}
	}
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Zoomed view of the \sn\ light curve.}  
		The full Kepler dataset spans 1{,}437 days (3.9 years), sampled at
		30 minute cadence.  The {\it top panel} shows the
		\texttt{PDCSAP} median-subtracted flux in units of
		parts-per-thousand ($\times 10^{-3}$).  The dominant signal
		is induced by starspots and plages.  The model for the
		stellar variability (orange line) is subtracted in the {\it
			bottom panel}, revealing the transits of \pn, as well as other
		deviations from the stellar variability model.  The 
		Figure Set available online shows the entire 3.9 years of
		available data.
		\label{fig:lightcurve}
	}
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f5.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Phase-folded transit of Kepler 1627b, with stellar
			variability removed.}  
    The 1-$\sigma$ model uncertainties and the maximum {\it a
    posteriori} model are shown as the faint purple band, and the dark
    purple line.  Gray points are individual \texttt{PDCSAP} flux
    measurements; black points are binned to 20 minutes, with a
    representative error bar shown.
    {\bf LGB FIXME: seems like the uncertainty is bigger than the RMS
    in the out of transit...}
		\label{fig:phasefold}
	}
\end{figure*}

Comparing our young star list against the Kepler field yielded two
discoveries.  The first, to be discussed in an upcoming analysis
by J. Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr
open cluster Theia\,520 (UBC\,1).  At least five Kepler planets are
known in the cluster, including the Kepler-52 and Kepler-968 systems.
The second discovery, and the focus of this work, is that Kepler
observed a portion of the $\delta$\,Lyr cluster, shown in
Figure~\ref{fig:skychart}.

While we have yet to perform an in-depth planet search of the {\bf
XXX} light curves Kepler acquired for the \cn, there is already a
known planet in the cluster: Kepler\,1627b.  We considered the
planet's previous statistical validation by \citet{morton_false_2016}
to be sufficient cause to examine the age of the star and its cluster more closely.
We collected the previously measured {\bf XXX} Kepler
rotation periods (CITEP MCQUILLAN2014), and measured {\bf YYY} new stellar
rotation periods for cluster members observed by TESS.
We also performed an isochronal analysis.
The results are shown in Figure~\ref{fig:age}.
From the ensemble perspective, which provides the strongest
constraints (SEE CITET SODERBLOM 2014),
the age of the star is $XX \pm YY$\,Myr.
Individual measurements of the star's rotation period and lithium
abundance corroborate this conclusion.

If the Kepler\,1627b transit signal truly corresponds to a planet,
this age would make it the youngest planet yet found by the main
Kepler mission.  A portion of the light curve is shown in
Figure~\ref{fig:lightcurve}, and the phase-folded transit is shown in
Figure~\ref{fig:phasefold}.  The dominant signal in the
\texttt{PDCSAP} photometry is a quasi-period starspot signal with a
peak-to-peak amplitude that varies between roughly 2\% and 8\%,
depending on the degree of asymmetry in spot coverage between the
stellar hemispheres.  The starspot signal repeats roughly every
2.606\,days, though this value varies by a few percent depending on
which portion of the light curve is analyzed {\bf (FIXME: do!)}.  The
details of how we model this signal to derive the parameters of the
planet are discussed in Section~\ref{app:gptransit}.  The results,
given in Table~\ref{tab:posterior}, are consistent with a mini-Neptune
sized planet ($3.56\pm 0.04\,R_\oplus$) on a circular 7.2\,day orbit,
around a G8V star ($0.91 \pm 0.05 R_\odot$) {\bf fixme: age}.

Could the transit signal be produced by anything other than a planet
orbiting this near-solar analog?  \citet{morton_false_2016} considered
the various false positive scenarios, including those of a foreground
eclipsing binary, a hierarchical triple-star binary system, and a
background eclipsing binary.  \citet{morton_false_2016} reported the
latter scenario to be the only possible case, with a (model-dependent)
probability of $\approx10^{-5}$.  While this statistical validation is
reassuring, ``validated" planets have been demonstrated to be false
positives \citep{shporer_three_2017}.  Most failures of the
statistical validation approach arise when attempting to separate the
smallest stars from identically-sized gas giant planets (CITE).  With a size
of $\approx 3.6\,R_\oplus$, Kepler\,1627b does not meet this
concern.  The fact that we detect a low-mass stellar companion in
archival Keck-NIRC2 Kp-band (2.12\,$\mu $m) imaging does not change
the conclusion: the companion contributes 1\% of the flux
observed by Kepler, which combined with the observed transit shape
cannot be invoked as an explanation for the observed data
(Section~\ref{app:companionstar}).

Kepler\,1627Ab therefore provides an interesting extremum in the ages of the
Kepler planets, and opens multiple avenues for detailed
study.  In order of ascending difficulty, observing its
Doppler transit ($\Delta v_{\rm RM} \approx 20$\ms) with HIRES or the
upcoming KPF spectrograph could yield a measurement of its orbital
obliquity (CITEP,CITEP).  A transmission
spectrum focused on either hydrogen or helium absorption could yield
insight into its atmosphere, which is expected to be outflowing at a
rate of {\bf X.X}\,$M_\oplus\,{\rm yr}^{-1}$ (CITE).  Finally, the
most challenging measurement that may currently be in reach would be a
measurement of its mass, through either detailed timing analyses of
the Kepler data, or through a high-cadence multi-color radial velocity campaign.  The mass would yield constraints on both the planet's
composition (do the mini-Neptunes form with gaseous envelopes?), as
well as its initial entropy (``hot'' or ``cold''? CITE).

Ultimately, our age measurement for Kepler\,6127Ab was enabled by
identifying its connection to the \cn\ through the Gaia data.  The
Supplementary Table~\ref{tab:v05} enables similar crossmatches to
already known planets, as well as forthcoming planets from the TESS
mission (CITE RickerJATIS2014).  An alternative approach also
likely to bear fruit is to identify kinematic associations
around exoplanet host stars using the Gaia data, and to then verify
these associations with ancilliary rotation periods and spectroscopy
(CITEP Tofflemire).

@@

@@

\section{Companion Star and False Positive Assessment}
\label{app:companionstar}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.6\textwidth]{f6.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Gaia EDR3 renormalized unit weight error (RUWE) point
    estimates for Kepler\,1627A and other members of
    the $\delta$\,Lyr cluster.}  Since other members of the cluster
    with similar colors have comparable degrees of photometric
    variability, the high RUWE estimate suggests that Kepler\,1627A is
    a binary. 
		\label{fig:ruwe}
	}
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.6\textwidth]{TEMP_binarity_ao.png}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Keck NIRC2 AO image of Kepler 1627A and Kepler 1627B.}  
    {\bf SCALEBAR DENOTES...}
    {\bf North is up, East is left}.
    \label{fig:ao}
	}
\end{figure*}

Our analysis of archival Keck-NIRC2 Kp-band (2.12\,$\mu $m) imaging
revealed the existence of a previously unreported stellar neighbor,
unresolved in the Gaia source catalog {\bf FIXME: uncertainties}.  The
NIRC2 images yield a projected separation $\rho=0\farcs17$, with
$\Delta {\rm Kp} = 2.5$.  Using the measured Gaia EDR3 parallax for
the system, this implies a projected separation of 54\,AU.  The
presence of this star is consistent with the excess noise in the Gaia
astrometric time-series (see Section~\ref{app:companionstar}).  Given
the low chance of a star imaged within {\bf X.X arcseconds} to be a
chance companion along the line of sight, we proceed under the
assumption that it is bound, and that the Kepler\,1627 system is
binary.

Unforuntately, we do not have any reliable color information about
Kepler\,1627B.  Based on the
tabulation\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
version \texttt{2021/03/02}.} by \citet{pecaut_mamajek_2013}, the
measured NIR-contrast for a main sequence G8V star corresponds to a
spectral type for the companion of $\approx$M2V ($M_\star \approx
0.44\,M_\odot$).  However, the companion should have a longer
pre-main-sequence contraction phase than the primary, which would
imply that this mass is overestimated by $\sim10$ to $20\%$.

Could the companion be creating a false positive signal?  The
companion star contributes $\approx$1\% of the total flux observed in
the Kepler aperture (0.7-2.0\% V-band to G-band dmag difference from
EEM table; todo is fix using model spectra).  The observed transit has
a depth of 0.18\%.  An 18\% deep eclipse of the secondary star would
therefore be needed to produce a deep enough signal.  The shape of the
observed signal requires allowed impact parameters to span 0.02 to
0.73 (Table~\ref{tab:posterior}); the body transiting the secondary
would therefore need to be non-grazing with $R_3/R_2 \approx 0.42$.
Assuming a $\approx 0.42R_\odot$ radius of the imaged secondary, this
would imply a tertiary stellar radius of $\approx 0.2R_\odot$.  This
scenario ultimately yields a contradiction, because it would require
an ingress and egress phase that each span $\approx$40\% of the
transit duration ($\approx 65\,{\rm minutes}$).  The actual measured
ingress and egress duration is $\approx 15\,{\rm minutes}$),
4.4$\times$ shorter.

\paragraph{Stellar density implied by transit duration}
The duration of the transit ($2.823 \pm 0.057 hr$) and the implied
stellar density ($2.31 \pm 0.52\,{\rm g\,cm}^{-3}$) could in theory
help rule between whether the transiting body orbits the primary or
secondary star.  Ultimately, our stellar density measurement is not
precise enough to render the blend scenario implausible.  At 30 Myr, a
0.40\,$M_\odot$ solar-metallicity dwarf is $\approx 26\%$ larger than
when it is fully contracted on the main sequence (0.40\,$R_\odot$
vs{.} 0.50\,$R_\odot$; CITEALT: Choi et al, MIST grids).  The
theoretically implied companion density of $3.2\,{\rm g\,cm}^{-3}$ is
indeed larger than the primary star's density of $1.80\,{\rm
g\,cm}^{-3}$ (measured through the HIRES reconaissance spectroscopy).
However the stellar density measured from the transit fitting ($2.31
\pm 0.52\,{\rm g\,cm}^{-3}$) is only discrepant at the
$\approx$2-$\sigma$ level from the M-dwarf blend scenario.  It is
instead the combination of the flux contrast, transit depth, and
ingress duration that rule out this scenario.
